Do investors in clean energy ETFs herd?

### 2024-10-10 alt_energy_herding.tex
added graphs. worked on cross-references

\author { 
Vassilios Babalos\footnote{Department of Accounting and Finance, University of Peloponnese, Kalamata,  Greece; Email: v.babalos@uop.gr.} \,\,
Xolani Sibande\footnote{Corresponding Author. Department of Economics, University of Pretoria, Pretoria, South Africa; Email: xolaniss@gmail.com. Economic Research Department, South African Reserve Bank, Pretoria, South Africa.} \,\, 
Riza Demirer\footnote{Department of Economics and Finance, Southern Illinois University Edwardsville, Edwardsville, IL 62026-1102, USA; Email: rdemire@siue.edu} \,\,
Rangan Gupta\footnote{Department of Economics, University of Pretoria, Pretoria, South Africa; Email: rangan.gupta@up.ac.za.}
}
\date{\today}
\maketitle

@@

The sample consists of alternative energy equity ETFs (green ETFs) that
are traded in the US markets. The number of available alternative energy
ETFs in our sample varied from 10 in the beginning of analysis to 30 at
the most. The period of analysis runs from May 1st of 2016 through 19th
June of 2024. The starting date was selected on the basis of the COP
Paris agreement. Daily logarithmic returns were computed from the
closing prices of ETFs for a total of 2122 observations.

@@

\begin{longtable}[l]{|p{3.89in}|p{1.99in}}
\caption{List of used alternative energy equity ETFs}\tabularnewline

\multicolumn{1}{!{\color[HTML]{FFFFFF}\vrule width 1pt}>{\raggedright}m{\dimexpr 3.89in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Arial}{ETF}}}} & \multicolumn{1}{!{\color[HTML]{FFFFFF}\vrule width 1pt}>{\centering}m{\dimexpr 1.99in+0\tabcolsep}!{\color[HTML]{FFFFFF}\vrule width 1pt}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Arial}{Source}}}} \\

\multicolumn{1}{!{\color[HTML]{FFFFFF}\vrule width 1pt}>{\raggedright}m{\dimexpr 3.89in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Arial}{ETF}}}} & \multicolumn{1}{!{\color[HTML]{FFFFFF}\vrule width 1pt}>{\centering}m{\dimexpr 1.99in+0\tabcolsep}!{\color[HTML]{FFFFFF}\vrule width 1pt}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Arial}{Source}}}} \\

\multicolumn{1}{!{\color[HTML]{FFFFFF}\vrule width 1pt}>{\raggedright}m{\dimexpr 3.89in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Arial}{ALPS\ CLEAN\ ENERGY\ ETF}}}} & \multicolumn{1}{!{\color[HTML]{FFFFFF}\vrule width 1pt}>{\centering}m{\dimexpr 1.99in+0\tabcolsep}!{\color[HTML]{FFFFFF}\vrule width 1pt}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Arial}{}}}} \\

\multicolumn{1}{!{\color[HTML]{FFFFFF}\vrule width 1pt}>{\raggedright}m{\dimexpr 3.89in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Arial}{BLUE\ HORIZON\ BNE\ ETF}}}} & \multicolumn{1}{!{\color[HTML]{FFFFFF}\vrule width 1pt}>{\centering}m{\dimexpr 1.99in+0\tabcolsep}!{\color[HTML]{FFFFFF}\vrule width 1pt}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Arial}{\ }}}} \\

\multicolumn{1}{!{\color[HTML]{FFFFFF}\vrule width 1pt}>{\raggedright}m{\dimexpr 3.89in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Arial}{SPDR\ S\&P\ KENSHO\ CLEAN\ POWER\ ETF}}}} & \multicolumn{1}{!{\color[HTML]{FFFFFF}\vrule width 1pt}>{\centering}m{\dimexpr 1.99in+0\tabcolsep}!{\color[HTML]{FFFFFF}\vrule width 1pt}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Arial}{\ }}}} \\

\multicolumn{1}{!{\color[HTML]{FFFFFF}\vrule width 1pt}>{\raggedright}m{\dimexpr 3.89in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Arial}{GLOBAL\ X\ CLEANTECH\ ETF}}}} & \multicolumn{1}{!{\color[HTML]{FFFFFF}\vrule width 1pt}>{\centering}m{\dimexpr 1.99in+0\tabcolsep}!{\color[HTML]{FFFFFF}\vrule width 1pt}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Arial}{\ }}}} \\

\multicolumn{1}{!{\color[HTML]{FFFFFF}\vrule width 1pt}>{\raggedright}m{\dimexpr 3.89in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Arial}{PROSHARES\ S\&P\ KENSHO\ CLEANTECH\ ETF}}}} & \multicolumn{1}{!{\color[HTML]{FFFFFF}\vrule width 1pt}>{\centering}m{\dimexpr 1.99in+0\tabcolsep}!{\color[HTML]{FFFFFF}\vrule width 1pt}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Arial}{\ }}}} \\

\multicolumn{1}{!{\color[HTML]{FFFFFF}\vrule width 1pt}>{\raggedright}m{\dimexpr 3.89in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Arial}{INVESCO\ MSCI\ SUSTAINABLE\ FUTURE\ ETF}}}} & \multicolumn{1}{!{\color[HTML]{FFFFFF}\vrule width 1pt}>{\centering}m{\dimexpr 1.99in+0\tabcolsep}!{\color[HTML]{FFFFFF}\vrule width 1pt}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Arial}{\ }}}} \\

\multicolumn{1}{!{\color[HTML]{FFFFFF}\vrule width 1pt}>{\raggedright}m{\dimexpr 3.89in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Arial}{FIRST\ TRUST\ GLOBAL\ WIND\ ENERGY\ ETF}}}} & \multicolumn{1}{!{\color[HTML]{FFFFFF}\vrule width 1pt}>{\centering}m{\dimexpr 1.99in+0\tabcolsep}!{\color[HTML]{FFFFFF}\vrule width 1pt}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Arial}{\ }}}} \\

\multicolumn{1}{!{\color[HTML]{FFFFFF}\vrule width 1pt}>{\raggedright}m{\dimexpr 3.89in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Arial}{FIDELITY\ CLEAN\ ENERGY\ ETF}}}} & \multicolumn{1}{!{\color[HTML]{FFFFFF}\vrule width 1pt}>{\centering}m{\dimexpr 1.99in+0\tabcolsep}!{\color[HTML]{FFFFFF}\vrule width 1pt}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Arial}{\ }}}} \\

\multicolumn{1}{!{\color[HTML]{FFFFFF}\vrule width 1pt}>{\raggedright}m{\dimexpr 3.89in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Arial}{GLDS.BLOOMBERG\ CN.\ EN.\ EQ.ETF}}}} & \multicolumn{1}{!{\color[HTML]{FFFFFF}\vrule width 1pt}>{\centering}m{\dimexpr 1.99in+0\tabcolsep}!{\color[HTML]{FFFFFF}\vrule width 1pt}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Arial}{\ }}}} \\

\multicolumn{1}{!{\color[HTML]{FFFFFF}\vrule width 1pt}>{\raggedright}m{\dimexpr 3.89in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Arial}{FST.NQ.CN.EDGE\ SMRT.GRID\ INFRA\ IDX\ ETF}}}} & \multicolumn{1}{!{\color[HTML]{FFFFFF}\vrule width 1pt}>{\centering}m{\dimexpr 1.99in+0\tabcolsep}!{\color[HTML]{FFFFFF}\vrule width 1pt}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Arial}{\ }}}} \\

\multicolumn{1}{!{\color[HTML]{FFFFFF}\vrule width 1pt}>{\raggedright}m{\dimexpr 3.89in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Arial}{DEFIANCE\ NEXT\ GEN\ H2\ ETF}}}} & \multicolumn{1}{!{\color[HTML]{FFFFFF}\vrule width 1pt}>{\centering}m{\dimexpr 1.99in+0\tabcolsep}!{\color[HTML]{FFFFFF}\vrule width 1pt}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Arial}{\ }}}} \\

\multicolumn{1}{!{\color[HTML]{FFFFFF}\vrule width 1pt}>{\raggedright}m{\dimexpr 3.89in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Arial}{DIREXION\ HYDROGEN\ ETF}}}} & \multicolumn{1}{!{\color[HTML]{FFFFFF}\vrule width 1pt}>{\centering}m{\dimexpr 1.99in+0\tabcolsep}!{\color[HTML]{FFFFFF}\vrule width 1pt}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Arial}{\ }}}} \\

\multicolumn{1}{!{\color[HTML]{FFFFFF}\vrule width 1pt}>{\raggedright}m{\dimexpr 3.89in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Arial}{GLOBAL\ X\ HYDROGEN\ ETF}}}} & \multicolumn{1}{!{\color[HTML]{FFFFFF}\vrule width 1pt}>{\centering}m{\dimexpr 1.99in+0\tabcolsep}!{\color[HTML]{FFFFFF}\vrule width 1pt}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Arial}{\ }}}} \\

\multicolumn{1}{!{\color[HTML]{FFFFFF}\vrule width 1pt}>{\raggedright}m{\dimexpr 3.89in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Arial}{ISHARES\ GLOBAL\ CLEAN\ EN.\ ETF}}}} & \multicolumn{1}{!{\color[HTML]{FFFFFF}\vrule width 1pt}>{\centering}m{\dimexpr 1.99in+0\tabcolsep}!{\color[HTML]{FFFFFF}\vrule width 1pt}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Arial}{\ }}}} \\

\multicolumn{1}{!{\color[HTML]{FFFFFF}\vrule width 1pt}>{\raggedright}m{\dimexpr 3.89in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Arial}{BLACKR.WLD.EXUS\ CRBN\ TSTN.READINESS}}}} & \multicolumn{1}{!{\color[HTML]{FFFFFF}\vrule width 1pt}>{\centering}m{\dimexpr 1.99in+0\tabcolsep}!{\color[HTML]{FFFFFF}\vrule width 1pt}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Arial}{\ }}}} \\

\multicolumn{1}{!{\color[HTML]{FFFFFF}\vrule width 1pt}>{\raggedright}m{\dimexpr 3.89in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Arial}{NUB.CBN.TSTN.\&\ INFRA}}}} & \multicolumn{1}{!{\color[HTML]{FFFFFF}\vrule width 1pt}>{\centering}m{\dimexpr 1.99in+0\tabcolsep}!{\color[HTML]{FFFFFF}\vrule width 1pt}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Arial}{\ }}}} \\

\multicolumn{1}{!{\color[HTML]{FFFFFF}\vrule width 1pt}>{\raggedright}m{\dimexpr 3.89in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Arial}{TCW\ TRANSFORM\ SYSTEMS\ ETF}}}} & \multicolumn{1}{!{\color[HTML]{FFFFFF}\vrule width 1pt}>{\centering}m{\dimexpr 1.99in+0\tabcolsep}!{\color[HTML]{FFFFFF}\vrule width 1pt}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Arial}{\ }}}} \\

\multicolumn{1}{!{\color[HTML]{FFFFFF}\vrule width 1pt}>{\raggedright}m{\dimexpr 3.89in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Arial}{VANECK\ URANIUM\ AND\ NUCLEAR\ ENERGY}}}} & \multicolumn{1}{!{\color[HTML]{FFFFFF}\vrule width 1pt}>{\centering}m{\dimexpr 1.99in+0\tabcolsep}!{\color[HTML]{FFFFFF}\vrule width 1pt}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Arial}{\ }}}} \\

\multicolumn{1}{!{\color[HTML]{FFFFFF}\vrule width 1pt}>{\raggedright}m{\dimexpr 3.89in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Arial}{NUVEEN\ GLOBAL\ NET\ ZERO\ TRANSITION\ ETF}}}} & \multicolumn{1}{!{\color[HTML]{FFFFFF}\vrule width 1pt}>{\centering}m{\dimexpr 1.99in+0\tabcolsep}!{\color[HTML]{FFFFFF}\vrule width 1pt}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Arial}{\ }}}} \\

\multicolumn{1}{!{\color[HTML]{FFFFFF}\vrule width 1pt}>{\raggedright}m{\dimexpr 3.89in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Arial}{SPDR\ MSCI\ USA\ CIM.\ PA.\ ALIGNED\ ETF}}}} & \multicolumn{1}{!{\color[HTML]{FFFFFF}\vrule width 1pt}>{\centering}m{\dimexpr 1.99in+0\tabcolsep}!{\color[HTML]{FFFFFF}\vrule width 1pt}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Arial}{\ }}}} \\

\multicolumn{1}{!{\color[HTML]{FFFFFF}\vrule width 1pt}>{\raggedright}m{\dimexpr 3.89in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Arial}{INVESCO\ GLOBAL\ CLEAN\ ENERGY\ ETF}}}} & \multicolumn{1}{!{\color[HTML]{FFFFFF}\vrule width 1pt}>{\centering}m{\dimexpr 1.99in+0\tabcolsep}!{\color[HTML]{FFFFFF}\vrule width 1pt}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Arial}{\ }}}} \\

\multicolumn{1}{!{\color[HTML]{FFFFFF}\vrule width 1pt}>{\raggedright}m{\dimexpr 3.89in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Arial}{FST.NQ.CN.EDGE\ GREY.ETF}}}} & \multicolumn{1}{!{\color[HTML]{FFFFFF}\vrule width 1pt}>{\centering}m{\dimexpr 1.99in+0\tabcolsep}!{\color[HTML]{FFFFFF}\vrule width 1pt}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Arial}{\ }}}} \\

\multicolumn{1}{!{\color[HTML]{FFFFFF}\vrule width 1pt}>{\raggedright}m{\dimexpr 3.89in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Arial}{GLOBAL\ X\ SOLAR\ ETF}}}} & \multicolumn{1}{!{\color[HTML]{FFFFFF}\vrule width 1pt}>{\centering}m{\dimexpr 1.99in+0\tabcolsep}!{\color[HTML]{FFFFFF}\vrule width 1pt}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Arial}{\ }}}} \\

\multicolumn{1}{!{\color[HTML]{FFFFFF}\vrule width 1pt}>{\raggedright}m{\dimexpr 3.89in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Arial}{GLOBAL\ X\ RENEWABLE\ ENERGY\ PRODUCERS}}}} & \multicolumn{1}{!{\color[HTML]{FFFFFF}\vrule width 1pt}>{\centering}m{\dimexpr 1.99in+0\tabcolsep}!{\color[HTML]{FFFFFF}\vrule width 1pt}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Arial}{\ }}}} \\

\multicolumn{1}{!{\color[HTML]{FFFFFF}\vrule width 1pt}>{\raggedright}m{\dimexpr 3.89in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Arial}{TRUESHARES\ EAG.GLB.\ RENWEN.ETF}}}} & \multicolumn{1}{!{\color[HTML]{FFFFFF}\vrule width 1pt}>{\centering}m{\dimexpr 1.99in+0\tabcolsep}!{\color[HTML]{FFFFFF}\vrule width 1pt}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Arial}{\ }}}} \\

\multicolumn{1}{!{\color[HTML]{FFFFFF}\vrule width 1pt}>{\raggedright}m{\dimexpr 3.89in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Arial}{VANECK\ LOW\ CARBON\ ENERGY\ ETF}}}} & \multicolumn{1}{!{\color[HTML]{FFFFFF}\vrule width 1pt}>{\centering}m{\dimexpr 1.99in+0\tabcolsep}!{\color[HTML]{FFFFFF}\vrule width 1pt}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Arial}{\ }}}} \\

\multicolumn{1}{!{\color[HTML]{FFFFFF}\vrule width 1pt}>{\raggedright}m{\dimexpr 3.89in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Arial}{SMARTETFS\ SUST.EN.\ II\ ETF}}}} & \multicolumn{1}{!{\color[HTML]{FFFFFF}\vrule width 1pt}>{\centering}m{\dimexpr 1.99in+0\tabcolsep}!{\color[HTML]{FFFFFF}\vrule width 1pt}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Arial}{\ }}}} \\

\multicolumn{1}{!{\color[HTML]{FFFFFF}\vrule width 1pt}>{\raggedright}m{\dimexpr 3.89in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Arial}{INVESCO\ SOLAR\ ETF}}}} & \multicolumn{1}{!{\color[HTML]{FFFFFF}\vrule width 1pt}>{\centering}m{\dimexpr 1.99in+0\tabcolsep}!{\color[HTML]{FFFFFF}\vrule width 1pt}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Arial}{\ }}}} \\

\multicolumn{1}{!{\color[HTML]{FFFFFF}\vrule width 1pt}>{\raggedright}m{\dimexpr 3.89in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Arial}{VIRTUS\ DUFF\ \&\ PHELPS\ CLEAN\ ENERGY\ ETF}}}} & \multicolumn{1}{!{\color[HTML]{FFFFFF}\vrule width 1pt}>{\centering}m{\dimexpr 1.99in+0\tabcolsep}!{\color[HTML]{FFFFFF}\vrule width 1pt}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Arial}{\ }}}} \\

\multicolumn{1}{!{\color[HTML]{FFFFFF}\vrule width 1pt}>{\raggedright}m{\dimexpr 3.89in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Arial}{GLOBAL\ X\ WIND\ ENERGY\ ETF}}}} & \multicolumn{1}{!{\color[HTML]{FFFFFF}\vrule width 1pt}>{\centering}m{\dimexpr 1.99in+0\tabcolsep}!{\color[HTML]{FFFFFF}\vrule width 1pt}}{\textcolor[HTML]{000000}{\fontsize{7}{7}\selectfont{\global\setmainfont{Arial}{\ }}}} \\

Results of baseline and rolling window analysis points to significant
herding for the whole period and in various instances.

Following the relevant literature we calculate the Cross Sectional
Absolute Deviation (CSAD) measure for each day in the following manner.
We compute the difference of the ith ETF and market return where market
return is proxied by the cross sectional average of returns for sample
of our ETFs available for each day using equation (1):

The behavior of the CSAD measure for US Alternative Energy ETFs is
presented in Figue 1.

At a later stage we estimate a non-linear regression as in
\citet{galariotis2015}.

The testing procedure of herding relies on the above Equation (2).
Rational asset pricing models predict a linear relationship between
return dispersion and market returns under normal conditions, a
relationship that is no longer valid in the presence of herding. Herding
behavior leads to an increasing or decreasing cross sectional dispersion
with respect to market returns. In other words, herding is captured by a
non-linear term in the standard pricing equation indicating a decreasing
or an increasing returns' dispersion.

@@

Stated differently, as \citet{chang2000} argue, in the case of herding
the coefficient on the non-linear term (γ2) will be negative and
statistically significant. Table 2 presens the results of herding for
the full sample employing the non-linear equation (2). The estimated
coefficient on market return is positive and highly significant as
expected.The estimated coefficient on the non-linear term is negative
(-1.2773) and statistically significant with a t-statistic of -9.71
suggesting that herd behavior is present and robust in the US
alternative energy ETFs.

@@

\begin{longtable}[l]{|p{1.99in}|p{1.99in}|p{1.99in}}
\caption{Estimation results of herding in the U.S. equity alternative energy ETFs}\tabularnewline

@@

There is ample evidence in the relevant literature that herding behavior
in various asset markets \citep[see][]{pochea2017} exhibits asymmetry
and time-varying characteristics. To this end, we proceed to estimate
Equation (2) using the quantile regression (QR) proposed by
\citet{koenker1978} and Table 3 presents the results of estimating
Equation (2) across various quantiles of the returns dispersion. Our
focus is on the herding coefficient γ2, as a significant negative value
of γ2 is indicative of herding. Such a finding is observed at two
quantiles namely 25\% and 50\% with a value of -1.1056 and -1.165 which
are highly significant. It is worth mentioning that the sign of the
herding coefficient remains negative for almost all quantiles while the
significance changes from significant to insifnificant while we move
from low and middle to upper quantiles (75 and 90\%).

@@

\begin{longtable}[l]{|p{1.39in}|p{1.39in}|p{1.39in}|p{1.39in}}
\caption{Estimation results of herding across various quantiles}\tabularnewline

@@

It is widely accepted that asset returns are characterized by asymmetry,
that is, return dispersion tend to behave differently in rising and
falling markets \citep[see][]{geert2000, zhou2013, longin2001}. Based on
the above and in order to provide additional insight on the herding
phenomenon we examine whether herding presents an asymmetric response on
days when the market is up vis-à-vis days when the market is down. To
this end, we augment Eq. 2 as follows:

\[
CSAD_t = \gamma_0 + \gamma_1(1-D) R_{m,t} + \gamma_2D R_{m,t} + \gamma_3(1-D) R_{m,t}  + \gamma_4D R_{m,t}^2 \epsilon_t
\] where D is a dummy variable that takes the value of 1 when the market
return is negative and 0 otherwise.

Therefore, our exploration of asymmetric behavior of herding phenomenon
is carried through the inspection of the statistical significance and
the sign of the two estimated coefficients γ3 vs γ4 (up vs down
markets). It should be noted, that examining the relationship between
returns dispersion and market-wide returns across various quantiles of
the returns distribution allows us to make more robust inference
regarding the true behavior of the phenomenon.

Table 4 reports the estimation results of herding in the up and down
markets based on Eq. 3. In general, we find that herding is more likely
to occur in down markets than in up markets, which is indicative of the
asymmetry of herding behavior.

@@

\begin{longtable}[l]{|p{0.99in}|p{0.99in}|p{0.99in}|p{0.99in}|p{0.99in}|p{0.99in}}
\caption{Estimation results of herding in up and down markets}\tabularnewline

@@

Herding is present at low quantiles when markets are rising with an
estimated coefficient γ3 of -1.3669 and -1.2383 and highly significant
respectively. However, when markets are declining, investors seem to
neglect their own information set and imitate the actions of others
resulting in a highly significant coefficient of herding ( γ4) across
four out of five quantiles. Furthermore, we find that in high quantiles
(75\% \& 90\%) and when markets are rising the coefficient of interest (
γ3) turns positive but insignificant.

@@

\begin{longtable}[l]{|p{3.78in}|p{2.14in}}
\caption{Estimation results of the probit model}\tabularnewline

@@

\begin{longtable}[l]{|p{2.23in}|p{1.90in}|p{1.86in}}
\caption{Estimation results of the probit model with high and low climate risk
indexes (above or below median)}\tabularnewline

@@

@@

>>>>

\author { 
Vassilios Babalos\footnote{Department of Accounting and Finance, University of Peloponnese, Kalamata,  Greece; Email: v.babalos@uop.gr.} \,\,
Xolani Sibande\footnote{Corresponding Author. Department of Economics, University of Pretoria, Pretoria, South Africa; Email: xolaniss@gmail.com. Economic Research Department, South African Reserve Bank, Pretoria, South Africa.} \,\, 
% Riza Demirer\footnote{Department of Economics and Finance, Southern Illinois University Edwardsville, Edwardsville, IL 62026-1102, USA; Email: rdemire@siue.edu} \,\,
Rangan Gupta\footnote{Department of Economics, University of Pretoria, Pretoria, South Africa; Email: rangan.gupta@up.ac.za.}
}
\date{\today}
\maketitle

@@

The sample consists of alternative energy equity ETFs (green ETFs) that
are traded in the US markets (see Table~\ref{tbl-data} in the
Appendix).\footnote{The data were sourced from
  \url{https://datastream.org/en-ca/}} The number of available
alternative energy ETFs in our sample varied from 10 in the beginning of
analysis to 30 at the most. The period of analysis runs from May
1\textsuperscript{st} of 2016 through 19\textsuperscript{th} June of
2024. The starting date was selected on the basis of the COP Paris
agreement. Daily logarithmic returns were computed from the closing
prices of ETFs for a total of 2122 observations.

Following the relevant literature we calculate the Cross Sectional
Absolute Deviation (CSAD) measure for each day in the following manner.
We compute the difference of the ith ETF and market return where market
return is proxied by the cross sectional average of returns for sample
of our ETFs available for each day using Equation~\ref{eq-csad}:

At a later stage we estimate a non-linear regression as in
\citet{galariotis2015}:

The testing procedure of herding relies on the above
Equation~\ref{eq-herding}. Rational asset pricing models predict a
linear relationship between return dispersion and market returns under
normal conditions, a relationship that is no longer valid in the
presence of herding. Herding behaviour leads to an increasing or
decreasing cross sectional dispersion with respect to market returns. In
other words, herding is captured by a non-linear term in the standard
pricing equation indicating a decreasing or an increasing returns'
dispersion.

Based on the above and in order to provide additional insight on the
herding phenomenon we examine whether herding presents an asymmetric
response on days when the market is up vis-à-vis days when the market is
down. To this end, we augment Equation~\ref{eq-herding} as follows:

\begin{equation}\phantomsection\label{eq-asymmetry}{
CSAD_t = \gamma_0 + \gamma_1(1-D) R_{m,t} + \gamma_2D R_{m,t} + \gamma_3(1-D) R_{m,t}  + \gamma_4D R_{m,t}^2 \epsilon_t
}\end{equation}

where \(D\) is a dummy variable that takes the value of 1 when the
market return is negative and 0 otherwise. Therefore, our exploration of
asymmetric behavior of herding phenomenon is carried through the
inspection of the statistical significance and the sign of the two
estimated coefficients \(\gamma_3\) vs \(\gamma_4\) (up vs down
markets).

@@

@@

The behaviour of the CSAD measure for US Alternative Energy ETFs is
presented in Figure~\ref{fig-csad}.

\begin{figure}[H]

\centering{

\includegraphics{alt_energy_herding_files/figure-pdf/fig-csad-1.pdf}

}

\caption{\label{fig-csad}Cross Sectional Absolute Deviation (CSAD) for
US Alternative Energy ETFs}

\end{figure}%

Stated differently, as \citet{chang2000} argue, in the case of herding
the coefficient on the non-linear term (\(\gamma_2\)) will be negative
and statistically significant. Table~\ref{tbl-herding} presents the
results of herding for the full sample employing the non-linear
Equation~\ref{eq-herding}. The estimated coefficient on market return is
positive and highly significant as expected.The estimated coefficient on
the non-linear term is negative (-1.2773) and statistically significant
with a t-statistic of -9.71 suggesting that herd behaviour is present
and robust in the US alternative energy ETFs.

@@

\begin{longtable}[l]{|p{1.99in}|p{1.99in}|p{1.99in}}

\caption{\label{tbl-herding}Estimation results of herding in the U.S.
equity alternative energy ETFs}

\tabularnewline

@@

There is ample evidence in the relevant literature that herding
behaviour in various asset markets \citep[see][]{pochea2017} exhibits
asymmetry and time-varying characteristics. To this end, we proceed to
estimate Equation~\ref{eq-herding} using the quantile regression (QR)
proposed by \citet{koenker1978} and Table~\ref{tbl-quantile_herding}
presents the results of estimating Equation~\ref{eq-herding} across
various quantiles of the returns dispersion. Our focus is on the herding
coefficient \(\gamma_2\), as a significant negative value of
\(\gamma_2\) is indicative of herding. Such a finding is observed at two
quantiles namely 25\% and 50\% with a value of -1.1056 and -1.165 which
are highly significant. It is worth mentioning that the sign of the
herding coefficient remains negative for almost all quantiles while the
significance changes from significant to insignificant while we move
from low and middle to upper quantiles (75\% and 90\%).

@@

\begin{longtable}[l]{|p{1.39in}|p{1.39in}|p{1.39in}|p{1.39in}}

\caption{\label{tbl-quantile_herding}Estimation results of herding
across various quantiles}

\tabularnewline

@@

It is widely accepted that asset returns are characterized by asymmetry,
that is, return dispersion tend to behave differently in rising and
falling markets \citep[see][]{geert2000, zhou2013, longin2001}. It
should be noted, that examining the relationship between returns
dispersion and market-wide returns across various quantiles of the
returns distribution allows us to make more robust inference regarding
the true behaviour of the phenomenon. Table~\ref{tbl-asymmetry_herding}
reports the estimation results of herding in the up and down markets
based on Equation~\ref{eq-asymmetry}. In general, we find that herding
is more likely to occur in down markets than in up markets, which is
indicative of the asymmetry of herding behaviour.

@@

\begin{longtable}[l]{|p{0.99in}|p{0.99in}|p{0.99in}|p{0.99in}|p{0.99in}|p{0.99in}}

\caption{\label{tbl-asymmetry_herding}Estimation results of herding in
up and down markets}

\tabularnewline

@@

Herding is present at low quantiles when markets are rising with an
estimated coefficient \(\gamma_3\) of -1.3669 and -1.2383 and highly
significant respectively. However, when markets are declining, investors
seem to neglect their own information set and imitate the actions of
others resulting in a highly significant coefficient of herding
(\(\gamma_4\)) across four out of five quantiles. Furthermore, we find
that in high quantiles (75\% and 90\%) and when markets are rising the
coefficient of interest (\(\gamma_3\)) turns positive but insignificant.

@@

\begin{longtable}[l]{|p{3.78in}|p{2.14in}}

\caption{\label{tbl-probit_analysis}Estimation results of the probit
model}

\tabularnewline

@@

\begin{longtable}[l]{|p{2.23in}|p{1.90in}|p{1.86in}}

\caption{\label{tbl-probit_analysis2}Estimation results of the probit
model with high and low climate risk indexes (above or below median)}

\tabularnewline

@@

Results of baseline and rolling window analysis points to significant
herding for the whole period and in various instances. \newpage

@@

\setcounter{section}{0}
\renewcommand{\thesection}{\Alph{section}}

\setcounter{table}{0}
\renewcommand{\thetable}{A\arabic{table}}

\setcounter{figure}{0}
\renewcommand{\thefigure}{A\arabic{figure}}
